---
title: "Class 11"
author: "Raidah Anisah Huda"
format: pdf
editor: visual
---

# Candy data

In today's class we will examine S3* Candy data and see if this helps us gain some more feeling for how PCA and other methods work.

```{r}
candy <- read.csv("candy-data.csv", row.names = 1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

There are 85 different candy types

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

There are 38 fruity candy types.

> Q. What are these fruity candy?

We can use the ==

```{r}
row.names(candy[candy$fruity == 1, ])
```

# How often does my favorite candy win

```{r}
candy["Twix", ]$winpercent

```

> Q3. What is your favorite candy in the dataset and what is it's winpercent value?

```{r}
candy["Swedish Fish", ]$winpercent

```

> Q4. What is the winpercent value for "Kit Kat"?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for "Tootsie Roll Snack Bars"?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
skimr::skim(candy)
library("skimr")
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, a majority of the columns are in a 0:1 scale but the `winpercent` column is on a different scal to the majoirty, with a scale of 0:100.

>Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero means the candy is not classified as containing chocolate. 

>Q8. Plot a histogram of winpercent values

In R basics plot
```{r}
hist(candy$winpercent)
```
```{r}
library(ggplot2)
hist<- ggplot(candy)+
  aes(winpercent)+
  geom_histogram()

hist
```


>Q9. Is the distribution of winpercent values symmetrical?

No, it appears that the winpercent values are skewed to the left had side of the histogram, indicating a skew towards lower winpercents.

>Q10. Is the center of the distribution above or below 50%?

It appears tht the center of the distribution is below 50% with a mean of :

```{r}
mean(candy$winpercent)
```

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

To answer this question I will need to:
-"subset" (a.k.a. "select", "filter") the candy dataset to just chocolate candy 
-get there winpercents
-calculate the mean of these. Then do the same for fruity candy an compare.

```{r}
#Filter/select/subset to just chocolate rows
choc.candy<- candy[as.logical(candy$chocolate),]
fruity.candy<- candy[as.logical(candy$fruity), ]

#Get there winpercent
choc.win <- choc.candy$winpercent
fruity.win <- fruity.candy$winpercent

#Calculate their mean winpercent value
mean(choc.win)
mean(fruity.win)
```
On average, the chocolate candy is higher ranked than the fruity candy.

>Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruity.win)
```
Yes, due to p-value being less than 5%, this difference between chocolate nad fruity candy is statistically significant.


# Overall Candy Rankings

There is a base R function called `sort()` for, guess what, sorting vecotrs of input.

```{r}
x<- c(5,2,10)

sort(x, decreasing = TRUE)
```
The buddy function to `sort()` that is often even more useful is called `order()`. It returns the "indices" of the input that would result in it being sorted.

```{r}
order(x)
x[order(x)]
```

>Q13. What are the five least liked candy types in this set?

I can order by the `winpercent`
```{r}
ord<- order(candy$winpercent)
head(candy[ord,],5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
ord.top<- order(candy$winpercent, decreasing = TRUE)
head(candy[ord.top,],5)
```
>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```
>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

```{r}
library(ggplot2)

my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "gold"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)

```

>Q17. What is the worst ranked chocolate candy?

Sixlets are the worst ranked chocolate candy

>Q18. What is the best ranked fruity candy?

Starburst are the best ranked fruity candy. 

# Taking a look at Price Percent

```{r}
my_cols[as.logical(candy$fruity)] = "red"
library(ggplot2)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent) +
  geom_point(col=my_cols)  

```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +
  geom_text()
```

To deal with overlapping labels I can use the **geom_repel** package

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +
  geom_text_repel(max.overlaps = 6, col=my_cols, size=3)
```


# Correlation Structure
Pearson correlation goes between -1 and +1 with zero indicating no correlation and values close to one being very highly correlated

```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruity and chocolate are anti-correlated

>Q23. Similarly, what two variables are most positively correlated?

Chocolate and Bars appear to be highly correlated as well as chocolate and winpercent


# PCA

The base R function for PCA is called `prcomp()` and we can set "scale=TRUE/FALSE"

```{r}
pca<- prcomp(candy, scale = TRUE)
summary(pca)
```
The main result fo PCA - i.e. the new PC plot (projection of candy on our new PC  axis in cotained in `pca$x
```{r}
pc<- as.data.frame(pca$x)
ggplot(pc)+
  aes(PC1, PC2, label=rownames(pc))+
  geom_point(col=my_cols)+
  geom_text_repel(max.overlaps = 6)
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus are picked up strongly by PC1 in the positive direction and this makes sense. 


