---
title: "Class 10: Structural Bioinformatics II"
author: "Raidah Anisah Huda"
format: pdf
editor: visual
---
#Comparative analysis of ADK

ADK (Adenelate Kinase) is an important drug target and we would love to know how it works - -i.e. molecular mechanism.
 
There has been lots of work done on this protein due to it's importance including lots of crystal structures. 

>Q10. Which of the packages above is found only on BioConductor and not CRAN? 

>Q11. Which of the above packages is not found on BioConductor or CRAN?: 

>Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 


We will begin with getting an example ADK seqeunce from the database. 
```{r}
library(bio3d)
aa <- get.seq("1ake_A")
aa
```

We can now run BLAST with this sequence

```{r}
#b<-blast.pdb(aa)
```
We can run `hits`
```{r}
#hits <- plot(b)
```

Lets wee whats in our `hits` object
```{r}
#hits$pdb.id

hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A')
```


Now we can download all these PDB structure files:

```{r}
files <- get.pdb(hits$pdb.id, path = "pdbs", split=TRUE, gzip=TRUE)
```



Now I want to align and supperpose these structures which are all over the place.
```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile= "msa")
```



Let's have a look at how our `pdbs` object:

```{r}
pdbs
```

Now we have our aligned and superposed structures we can perform all sorts of analysis on them. Let's do PCA.

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```
Presults of PCA an Adenlyat kinase X-ray structures. Each dot represents one PDB structures,

We can cluster the structure by RMSD (or any other method).

```{r}
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps.rd)
```

We can make a wee movie - also called a trajectory of the major differences (i.e. structural displacments) od ADK.

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```







